\LARGE \bf Multi-dimensional Optimization of Gasoline-fueled Variable Pitch Multirotor Aircraft

### 2016-04-12 paper.tex
worked on Engine models, background, and various other little things

Incorporating a gas powered engine into a multirotor requires more then just exchanging the electric motors and batteries for an engine and a gas tank respectively. The control of such a vehicle would also fundamentally change.  Electric multirotors are controlled by spinning the rotors faster at slower to produce varying thrust and torque. Gad engines are unable to rev up as fast as electric motors meaning a fixed speed and variable pitch propeller is better suited for a gas powered vehicle.

@@

There are many commercially available parts for a gas quad, including engines and variable pitch helicopter propellers.

@@

Because the quad is going to be build using commercially available off-the-shelf parts the optimization must be closely based on reality.  This means we only can pick the size of engine that we are going to use.  All the other data about the engine (mass, maximum power, etc.) have to be a function of displacement size.  We collected data about popular two stroke RC aircraft engines from several sources to produce the models. One source was Desert Aircraft \cite{da2016}.

\begin{figurehere}
	\includegraphics[width=0.5\textwidth]{mass.png}
	\caption{The empirical based surrogate model of engine mass as a function of displacement size. Data from desertaircraft.com \cite{da2016}}
		\label{fig:mass}
\end{figurehere}

\begin{figurehere}
	\includegraphics[width=0.5\textwidth]{max_power.png}
	\caption{Empirical based surrogate model of engine maximum power as a function of displacement size. Data from desertaircraft.com \cite{da2016} and barnardmicrosystems.com \cite{barnardmiro2016}}
		\label{fig:power}
\end{figurehere}

The engine efficiency (friction and heat transfer losses), however, is a function of power capability of the engine (max power) and the power engine is required to produce.  For this we will need a better model.  \textit{We are still working on this. Since efficiency numbers for RC engines aren't publicly available, we need a scientific model.  For now we just assume a constant efficiency. We plan to work on this in the future.}

Although the off-the-shelf constraint also applies to choosing the rotor parameters, continuous variables approximating the rotor specs are used and then analyzed as follows.

>>>>

Incorporating a gas powered engine into a multirotor requires more then just exchanging the electric motors and batteries for an engine and a gas tank respectively. The control of such a vehicle would also fundamentally change.  Electric multirotors are controlled by spinning the rotors faster at slower to produce varying thrust and torque. Gas engines are unable to rev up as fast as electric motors meaning a fixed speed and variable pitch propeller is better suited for a gas powered vehicle.

@@

There are many commercially available parts for a gas quad, including engines \cite{da2016} and variable pitch helicopter propellers \cite{align2016}. However a gas powered multirotor would required a custom frame for mounting the engine and supplying power from the engine to the individual rotors.

FAA requirements...

@@

Because the quad is going to be build using commercially available off-the-shelf parts the optimization must be closely based on reality.  The analysis must, therefore, either be based on observations from available hardware or attempt analyze the hardware in its available configuration.  The following sections present some surrogate models for approximating engine parameters as well as aerodynamic analysis of common variable pitch RC propellers.

\subsection{Engine}

The designer essentially can only choose the size of engine. All the other data about the engine (mass, maximum power, etc.) have to be a function of displacement size.  Data was collected for popular two stroke RC aircraft engines from several sources to produce the models for other unknown parameters. One source was Desert Aircraft \cite{da2016}.  Figure~\ref{fig:mass} shows the collected data for engine size and mass.  We performed a linear least squares bast fit to produce a model for use in our analysis.

\begin{figure}
	\includegraphics[width=0.5\textwidth]{mass.png}
	\caption{The empirical based surrogate model of engine mass as a function of displacement size. Data from desertaircraft.com \cite{da2016}. $\ y = ax + b \quad a = 0.02145 \quad b = 0.25412$}
		\label{fig:mass}
\end{figure}

Data was also used to produce a model of the power output of the engine as a function of size. Figure~\ref{fig:power} shows that collected data for engine size and power.  Again we initially used a linear least squares fit to the data as well as compared our fit to data produced by Barnard Microsystems \cite{barnardmiro2016}. However both models were inadequate for our analysis because they contain a non-zero y-intercept.  A non-zero y-intercept would allow an infinitely small engine to produce power and thus would allow the aircraft to fly infinitely. To fix this deficiency we applied a quadratic fit to the data and forced it to do through the origin.

\begin{figure}
	\includegraphics[width=0.5\textwidth]{max_power.png}
	\caption{Empirical based surrogate model of engine maximum power as a function of displacement size. Notice the linear models have a non-zero y intercept but the quadratic fit goes through the origin. Data from desertaircraft.com \cite{da2016} and barnardmicrosystems.com \cite{barnardmiro2016} $\ y = ax^2 + bx \quad a = -0.04548 \quad b = 79.504$}
		\label{fig:power}
\end{figure}

The engine efficiency was not modeled for this project and was assumed a constant parameter.  Within the analysis the shaft output power of the engine was 15\% of the power consumed from the fuel.
